---
author: Di Shao
categories:
- communication
date: "2021-01-02"
draft: false
excerpt: Make publication-ready summary table.
layout: single
subtitle:
title: a learning notes of gtsummary 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE}
# install.packages("gtsummary")
library(gtsummary)
library(tidyverse)
```

## Basic Usage

You can also embed plots, for example:

```{r warning=FALSE}
head(trial)
trial2 <- trial %>% select(trt, age, grade)
trial2 %>% tbl_summary()
```
## split by treatment group and add p value
```{r warning=FALSE}
trial2 %>% tbl_summary(by = trt) %>% add_p()
```

## Customize Output

### Modifying tbl_summary() function arguments

```{r warning=FALSE}
trial2 %>%
  tbl_summary(
    by = trt,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} / {N} ({p}%)"),
    digits = all_continuous() ~ 2,
    label = grade ~ "Tumor Grade",
    missing_text = "(Missing)"
  )
```

### {gtsummary} functions to add information
add_n, add_p, add_overall, add_difference, add_q

### {gtsummary} functions to format table
```{r warning=FALSE}
trial2 %>%
  tbl_summary(by = trt) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>%
  add_overall() %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Treatment Received**") %>%
  modify_footnote(
    all_stat_cols() ~ "Median (IQR) or Frequency (%)"
  ) %>%
  bold_labels()
```

### {gt} functions to format table

```{r warning=FALSE}
trial2 %>%
  tbl_summary(by = trt, missing = "no") %>%
  add_n() %>%
  as_gt() %>%
  gt::tab_source_note(gt::md("*This data is simulated*"))
```

### Select Helpers
All {tidyselect} helpers available throughout the tidyverse
Additional {gtsummary} selectors: all_continuous() all_categorical()

### Multi-line Continuous Summaries
```{r warning=FALSE}
trial2 %>%
  select(age, trt) %>%
  tbl_summary(
    by = trt,
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c("{N_nonmiss}",
                                     "{median} ({p25}, {p75})", 
                                     "{min}, {max}"),
    missing = "no"
  ) %>%
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2))
```

## Advanced Customization

```{r warning=FALSE}
tbl_summary(trial2, by = trt) %>%
  as_gt(include = -cols_align) %>%
  gt::tab_source_note(gt::md("*This data is simulated*"))
```

## Cross Tables

```{r warning=FALSE}
trial %>%
  tbl_cross(
    row = stage,
    col = trt,
    percent = "cell"
  ) %>%
  add_p()
```

## set theme
```{r}
theme_gtsummary_journal(journal = "nejm")
# Setting theme `New England Journal of Medicine`
trial %>%
  select(trt, age, grade) %>%
  tbl_summary(by = trt) %>%
  add_p()
```



